---
title: "SRPartitioning"
author: "Jinshi"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, echo = FALSE,
                      fig.height = 5, fig.width = 7)

DATA_DIR <- 'Data'
OUT_DIR <- 'outputs'
```


```{r loading package}
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
theme_set(theme_bw())
library(here)
library(hexbin)
library(cowplot)
library(SpatialEpi)
library(patchwork)
library(ggridges)
# Note raster masks 'tidyr::extract' and 'dplyr::select'
library(raster)
library(randomForest)
library(caret)
library(e1071)
library(car) # for unbalanced anova
library(pdp)
source("functions.R")
```


```{r prepare data}
srdb_all <- drake::readd(srdb_all) %>% 
  mutate(MAT = coalesce(MAT, MAT_WC),
         MAP = coalesce(MAP, MAP_WC),
         absLatitude = abs(Latitude),
         MiddleClimate = as.factor(MiddleClimate)) %>% 
  filter(RC_annual < 1 & RC_annual > 0)

# EVI issue: /1000 and take care of < 0 issue
srdb_all$EVI_mean = srdb_all$EVI_mean / 1000
srdb_all$EVI_mean %>% min(na.rm = TRUE)

# ecosystem2 NA to NotReported
srdb_all$Ecosystem2 <- as.character(srdb_all$Ecosystem2)
srdb_all[is.na(srdb_all$Ecosystem2),]$Ecosystem2 <- "Not reported"
# Make sure there is no missing data
allData = srdb_all %>% 
  dplyr::select(RC_annual, MAT, MAP, AM_percent, EM_percent, ER_percent, NM_percent,
                BM_aboveground, BM_belowground, N_dep_half_deg, N_dep_1993,
                MiddleClimate, gpp_fluxcom,
                Ecosystem2, Latitude, Longitude,
                bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI_mean,
                Partition_method) %>% 
  na.omit()

allData$Ecosystem2 <- as.factor(allData$Ecosystem2)
  
# Extract global climate space data
map_mat_global <- drake::readd(map_mat_global)

# prepare data for global RC_annual prediction
globalData <- drake::readd(globalData) %>% 
  na.omit() %>% 
  mutate(absLatitude = abs(Latitude)) ->
  globalData

# EVI issue: /1000 and take care of < 0 issue
globalData$EVI_mean = globalData$EVI_mean / 1000
globalData$EVI_mean[globalData$EVI_mean < 0] = 0
globalData$EVI_mean %>% min(na.rm = TRUE)

counties <- drake::readd(counties)
```



```{r check NULL values}
srdb_all %>% 
  filter(is.na(MAP))

allData %>% dplyr::select(Ecosystem2) %>% unique()

srdb_all %>% 
  ggplot(aes(N_dep_1993, N_dep_half_deg)) + 
  geom_point()
```


```{r srdb vasialization}
ggplot(srdb_all, aes(x = Leaf_habit, fill = Leaf_habit)) +
  geom_bar() +
  theme_minimal()

srdb_all %>% 
  ggplot(aes(Rs_annual, Leaf_habit, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(Annnual~R[S]~(g~C~m^{-2}~yr^{-1}))) +
  ylab(element_blank()) +
  scale_y_discrete(labels= c("Not available", "Deciduous", "Evergreen", "Mixed"))

srdb_all %>% 
  ggplot(aes(RC_annual, Leaf_habit, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[C]~Annual)) +
  ylab(element_blank()) +
  scale_y_discrete(labels= c("Not available", "Deciduous", "Evergreen", "Mixed"))

```


## Pulling the Climate Data and Comparing the Distribution to that in the SRDB
The first is a plot of the global distribution of mean annual temperature (MAT) and precipitation (MAP). 
The second also includes the values from all SRDB sites, plotted on top of the first graph.
```{r Site map, cache = TRUE, fig.height = 7, fig.width = 8}
# Plot panel a - global map with RC_annualsites
ggplot(data = counties) + 
  geom_polygon(aes(x = long, y = lat, group = group),
               color = "white", fill = 'gray', alpha = 0.5) + 
  guides(fill=FALSE) +
  geom_point(data = allData,
             aes(x=Longitude, y=Latitude, col = Ecosystem2),
             alpha = 0.75,
             stroke = 0.75 ) +
  scale_x_continuous(name="Longitude", breaks=seq(-180,180, 60),labels = seq(-180,180, 60))+
  scale_y_continuous(limits = c(-60, 80),name="Latitude", breaks=seq(-60,80,30),labels = seq(-60,80,30)) +
  scale_size_continuous(name = "RC_annual") +
  scale_colour_discrete(name = element_blank(),
                        breaks=c("Agriculture", "Grassland", "Not reported", "Savanna", "Shrubland", "T&B FOR", "Tropic FOR", "Urban", "Wetland"),
                        labels=c("Agriculture (n=261)", "Grassland (n=140)", "Not reported (n=2)", "Savanna (n=71)", "Shrubland (n=21)",
                                 "T&B FOR (n=256)", "Tropic FOR (n=70)", "Urban (n=2)", "Wetland (n=59)")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) ->
  gmap_a

# This chunk pulls climate data from WorldClim2 and makes some graphs from the mean annual temperature and precipitation data
# Download worldclim data for precip and tmean if necessary, into w10/ folder
# global distribution of MAT and MAP with points from the SRDB sites
map_mat_global %>% 
  ggplot(aes(x = mat/10, y = map)) +
  geom_hex(bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  geom_point(data = srdb_all, aes(x = MAT_WC, y = MAP_WC),
             color = "black", shape = 1, size = 1.5, na.rm = TRUE) +
  # theme(legend.position = c(0.2, 0.65)) +
  labs(x = "MAT (°C)", y = "MAP (mm)") ->
  gmap_b

gmap_a / gmap_b +
  plot_annotation(tag_levels = 'a')

allData %>% 
  count(Ecosystem2)

srdb_all %>% 
  count(Ecosystem2)
```


```{r check WorldClim2 data}
# Comparing MAT
ggplot(srdb_all, aes(x = MAT_WC, y = MAT)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MATs to SRDB MATs", x = "WorldClim2 MAT (°C)", y = "Reported SRDB MAT (°C)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)

# Comparing MAP
ggplot(srdb_all, aes(x = MAP_WC, y = MAP)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MAPs to SRDB MAPs", x = "WorldClim2 MAP (mm)", y = "Reported SRDB MAP (mm)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)

## Makes some scatterplots with the SRDB data
ggplot(srdb_all, aes(x = MAT_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()

ggplot(srdb_all, aes(x = MAP_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAP_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = MAT_WC, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = MAP_WC, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = MAP_WC, col = Leaf_habit)) +
  geom_point() +
  theme_minimal()
```

```{r Micro data graphs}

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = AM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = EM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = ER_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = NM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

# None of these seem to show correlations with mean annual precipitation, with the exception of ER mycorrhizae.
ggplot(srdb_all, aes(x = MAT_WC, y = AM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = EM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = NM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

# Microriza vs MAP
ggplot(srdb_all, aes(x = MAP_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = AM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = EM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = ER_percent, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = NM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()
```


## RF modelling
### Improving the model
Here is a graph of the error vs the number of trees in the model and a graph of how the error varies with different values of mtry.
mtry is a parameter for this random forest function that sets how many variables are tried when making each node of a regression tree.
The number of trees and this mtry value are both input parameters used to produce the random forest model.
```{r RF determine number of parameters}
# Tune with different mtry values
set.seed(20201217)
tuneRF(subset(allData, select = c(MAT, MAP, AM_percent, EM_percent, ER_percent, NM_percent,
                                   BM_aboveground, BM_belowground, N_dep_half_deg, bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI_mean)),
       allData$RC_annual,
       stepFactor = 3,
       ntreetry = 200, # Based on the plot above, 200 trees seems reasonable
       trace = TRUE,
       improve = 1e-5) ->
  tune_rf

as.data.frame(tune_rf) %>% 
  mutate(OOBError = round(OOBError, 5)) %>% 
  ggplot(aes(x = mtry, y = OOBError)) +
  geom_point() +
  geom_line() +
  xlab (expression(m[try])) +
  ylab ("OOB Error") 
```

First, I will use the default settings in the randomForest package to make a random forest model to predict the root contribution to Rs (RC). 
```{r create RF model for prediction}
set.seed(20201007)
## Use random forest machine learning to create a model to predict Rs partitioning
# Set seed to a specific value, for reproducability
# Make a random forest model
rf <- randomForest(RC_annual ~ MAT + MAP +
                     AM_percent + EM_percent + ER_percent + NM_percent + BM_aboveground + BM_belowground +
                     N_dep_half_deg + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI_mean +
                     MiddleClimate, 
                   data = allData,
                   # %>% sample_frac(frac_samp)
                   importance = TRUE, 
                   proximity = TRUE,
                   ntree = 200,
                   mtry = 4,
                   na.action = na.exclude)   

plot(rf)
importance(rf)
varImpPlot(rf)
```

## SRDB RC_annual prediction
```{r SRDB RC_annual prediction check}
## Investigate how well the model works at predicting training data RC
# Use this model to predict RC values for the training data and put it all in one dataframe

# Plot training data RC values and the predicted values from the model
allData %>% 
  mutate(prediction = predict(rf),
         residuals = RC_annual - prediction) ->
  allData

md_lm <- lm(RC_annual ~ prediction, data = allData)
summary(md_lm)$r.squared %>% round(2) -> r2

allData %>% 
  ggplot(aes(y = RC_annual, x = prediction)) +
  lims(x = c(0, 1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1, color = "red") +
  labs(y = expression(Measured~R[root]~contribution~to~R[S]),
       x = expression(Predicted~R[root]~contribution~to~R[S])) +
  annotate("text", x = 0.1, y = 0.9, label = paste0("italic(R) ^ 2 == ", r2), parse = TRUE) +
  scale_y_continuous(position = "left") ->
  srdb_pred_vs_measure

srdb_pred_vs_measure

allData %>% 
  ggplot(aes(x = prediction, y = residuals)) +
  lims(x = c(0, 1)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed", size = 1, color = "red")

```


```{r RF diagnose, fig.width=8, fig.height=6}
# panel a
qplot(1:200, rf$mse ) + geom_line() + theme_bw() +
  xlab ("Number of trees (n)") + ylab("MSE") +
  theme(axis.title.y = element_text(angle = 90, hjust = 0.5) ) ->
  p_a

# panel b
tibble(x = c("ErM%", "NM%", "EcM%", "Climate", "AM%", "BMa", "N dep", "Clay%", "MAT", "SOC", "BMb", "EVI", "BD", "MAP"),
       y = sort(importance(rf, type = 1)[1:15])) %>% 
  ggplot(aes(x = reorder(x, y), y)) +
  geom_bar(stat = "identity", width = 0.75) +
  coord_flip() +
  ylab ("Change of MSE (%)") +
  theme(axis.title.y = element_blank(),
        panel.grid.minor = element_blank()) ->
  rf_b

# panel c
tibble(x = c("Erm%", "NM%", "EcM%", "Climate", "AM%", "Clay%", "N dep", "SOC", "BMb", "BMa", "MAT", "EVI", "BD", "MAP"),
       y = sort(importance(rf, type = 2)[1:15])) %>% 
  ggplot(aes(x = reorder(x, y), y)) +
  geom_bar(stat = "identity", width = 0.75) +
  coord_flip() +
  ylab ("Change of node purity") +
  theme(axis.title.y = element_blank(),
        panel.grid.minor = element_blank()) ->
  rf_c

p_a / (rf_b | rf_c) + plot_annotation(tag_levels = 'a')
```


## Global RC_annual prediction
```{r Predict global RC_annual}
levels(globalData$Ecosystem2)
levels(allData$Ecosystem2)

levels(globalData$MiddleClimate)
levels(allData$MiddleClimate)
# Predict global RC_annual
globalData$RC_predict = predict(rf, globalData)

# Make the default theme have a centered title
theme_update(plot.title = element_text(hjust = 0.5))

# Plot predictions globally
globalData %>% 
  ggplot(aes(Longitude, Latitude, color = RC_predict)) + 
  geom_point(size = 0.3) +
  lims(x = c(-180, 180), y = c(-60, 90)) +
  labs(color = "RC prediction") +
  scale_color_viridis_c()

# Look at the distribution of predictions
globalData %>% 
  ggplot(aes(x = RC_predict)) +
  geom_histogram(color = "black", fill = "lightgrey") +
  labs(x = "RC Prediction")

# get all prediction and calculate confidence interval
all_trees_pred <- predict(rf, globalData, predict.all=TRUE)
all_trees_RC <- all_trees_pred$individual
globalData$rc_sd <- apply(all_trees_RC, 1, sd)
globalData$rc_cv <- round(globalData$rc_sd/globalData$RC_predict *100, 2)

# outputs RC_prediction
writ_file(globalData %>%
            dplyr::select(Latitude, Longitude, RC_predict, rc_sd, rc_cv),
          "Global_RC_annual_prediction.csv")

writ_file(globalData,
          "Global_RC_annual_prediction2.csv")

globalData$RC_predict %>% mean()
globalData$RC_predict %>% sd()
```


```{r check N_dep}
ggplot(globalData, aes(Longitude, Latitude, color = EVI_mean)) + 
  geom_point(size = 0.3) +
  lims(x = c(-180, 180), y = c(-90, 90)) +
  labs(color = "RC prediction") +
  scale_color_viridis_c()

ggplot(globalData, aes(Longitude, Latitude, color = Ecosystem2)) + 
  geom_point(size = 0.3) +
  lims(x = c(-180, 180), y = c(-90, 90)) +
  labs(color = "RC prediction") 

# N_dep_half_deg

# MAT + MAP + AM_percent + EM_percent + ER_percent + NM_percent + BM_aboveground + BM_belowground +
#   N_dep_half_deg + MiddleClimate + Ecosystem2 + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI_mean
```



```{r RC_annual vs RC_prediction global, fig.width=8, fig.height=7}
globalData$RC_annual <- globalData$RC_predict

bind_rows(
  globalData %>% 
    mutate(Type = "Global") %>% 
    dplyr::select(RC_annual, Type),
  
  allData %>% 
    mutate(Type = "SRDB") %>% 
    dplyr::select(RC_annual, Type)  ) %>% 
  ggplot(aes(RC_annual, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[root]~contribution~to~R[S])) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0)) +
  scale_y_discrete(labels= c("Global", "SRDB"),
                   position = "left") ->
  overall_comparisom

globalData$RC_annual %>% min()
globalData$RC_annual %>% max()
globalData$RC_annual %>% mean()

allData$RC_annual %>% min()
allData$RC_annual %>% max()
allData$RC_annual %>% mean()

srdb_pred_vs_measure / overall_comparisom +  plot_annotation(tag_levels = 'a')
```

```{r}
# do local corelation and get coefficiency
# corLocal(x, y, ngb=5, 
#      method=c("pearson", "kendall", "spearman"), test=FALSE, filename='', ...)
```


## SI
```{r Figure S1, fig.height=8, fig.width=6}
# Unbalanced ANOVA
allData$Partition_method %>% unique()
allData %>% 
  mutate(Partition_group = case_when(
    Partition_method %in% c("Comparison", "Exclusion", "Girdling") ~ "Exclusion",
    Partition_method %in% c("Isotope") ~ "Isotope",
    Partition_method %in% c("Extraction") ~ "Physical seperation",
    Partition_method %in% c("Model", "TBCA", "Subtraction") ~ "Component",
    TRUE ~ "Other")) ->
  allData

allData %>% 
  count(Partition_group)

Anova(lm(RC_annual ~ Partition_group, data = allData), type="III") -> anov_partition
summary(anov_partition)
Anova(lm(RC_annual ~ MiddleClimate, data = allData), type="III") -> anov_climate
summary(anov_climate)

allData %>% 
  dplyr::select(RC_annual, Partition_group) %>% 
  na.omit() %>% 
  filter(RC_annual <= 1 & RC_annual >= 0) %>% 
  # filter(Partition_method %!in% c("", "Other", "Girdling", "TBCA")) %>% #filter those out (not too much data)
  ggplot(aes(Partition_group, RC_annual, fill = Partition_group)) +
  geom_violin(draw_quantiles = c(.25, 0.5, .75)) +
  theme(legend.position = "none",
        # axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 0, vjust = .5)) +
  scale_x_discrete(breaks = c("Component", "Exclusion", "Isotope", "Physical seperation", "Other"),
                   labels = c("Component", "Exclusion", "Isotope", "Physical seperation", "Other") ) +
  labs(y = expression(R[root]~Contribution~to~R[S]),
       x = " ") -> plot1

allData %>% 
  filter(!is.na(MiddleClimate)) %>%
  ggplot(aes(x = MiddleClimate, y = RC_annual, fill = MiddleClimate)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 15, vjust = .5)) +
  scale_x_discrete(labels = c("Tropic", "Arid", "Temperate humid", "Temperate summer dry", "Temperate winter dry", 
                              "Boreal humid", "Boreal dry", "Arctic")) +
  labs(y = expression(R[root]~Contribution~to~R[S]),
       x = element_blank()) -> plot2

plot1 / plot2
```

# test factors affect RC
```{r RC vs Rs_annual}
# load Bond-Lamberty 2004 data
bond2004 <- read_file('2004_paper_data.csv')
# RC model
allData2 = srdb_all %>% 
  dplyr::select(Rs_annual, RC_annual, MAT, MAP, AM_percent, EM_percent, ER_percent, NM_percent,
                BM_aboveground, BM_belowground, N_dep_half_deg, N_dep_1993,
                MiddleClimate, gpp_fluxcom,
                Ecosystem2, Ecosystem_type, Latitude, Longitude,
                bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI_mean,
                Partition_method) %>% 
  na.omit()

RC_v5_mod <- lm(RC_annual ~ log(Rs_annual), data = allData2)
RC_v5_mod2 <- lm(RC_annual ~ log(Rs_annual), data = allData2 %>% filter(Rs_annual < 2500))
summary(RC_v5_mod)
summary(RC_v5_mod2)
allData2$Ecosystem2 %>% unique()
allData2$Ecosystem_type %>% unique()
RC_v5_for <- lm(RC_annual ~ log(Rs_annual), data = allData2 %>% filter(Ecosystem_type %in% c("Forest")))
summary(RC_v5_for)

bind_rows(
  allData2 %>% 
    dplyr::select(Rs_annual, RC_annual) %>% 
    mutate(Source = "SRDB-V5"),
  
  bond2004 %>% 
    dplyr::select(Rs, RC) %>% 
    rename(Rs_annual = Rs, RC_annual = RC) %>% 
    mutate(Source = "Old")) ->
  allData3

allData3 %>% 
  ggplot(aes(Rs_annual, RC_annual, col = Source)) +
  geom_point(alpha = 0.5) +
  geom_point(aes(Rs, RC), data = bond2004, col = "blue") +
  # ggtitle(expression("RC=-0.07+0.08ln(Rs), R"^2*"=0.02, n=478")) +
  stat_function(fun = function(x) { -0.66 + 0.16*log(x) }, col = 'blue', size = 1.5) +
  stat_function(fun = function(x) { 0.082 + 0.05*log(x) }, col = 'gray', size = 1.5) +
  stat_function(fun = function(x) { 0.1807 + 0.03427*log(x) }, col = 'green', size = 1.5) + # forest
  stat_function(fun = function(x) { 0.155 + 0.03876*log(x) }, col = 'black', size = 1.5) +
  labs(x = element_blank(), 
       y = expression(R[root]~contribution~to~R[S]~(RC))) +
  scale_color_manual(name = "Data",
                     values = c("blue", "black"),
                     breaks = c("Old", "SRDB-V5"),
                     labels = c("Old", "New")) +
  geom_segment(aes(x = 2500, y = 0.95, xend = 2500, yend = 0.65),
               arrow = arrow(length = unit(0.25, "cm")),
               size = 0.5,
               col = "blue") +
  geom_segment(aes(x = 2500, y = 0.125, xend = 2500, yend = 0.43),
               arrow = arrow(length = unit(0.25, "cm")),
               size = 0.5,
               col = "black") +
  annotate("text", x = 2600, y = 0.95, label = paste0("RC = -0.66 + 0.16 ln(Rs)"), col = "blue", hjust = 0) +
  annotate("text", x = 2600, y = 0.85, label = paste0("italic(R) ^ 2 == 0.38 (n == 63)"), parse = TRUE, col = "blue", hjust = 0) +
  annotate(geom="text", x=2600, y=0.75, label="Bond-Lamberty (2004)", color="blue", hjust = 0) +
  
  annotate("text", x = 2600, y = 0.35, label = paste0("RC = 0.16 + 0.04 ln(Rs)"), col = "black", hjust = 0) +
  annotate("text", x = 2600, y = 0.25, label = paste0("italic(R) ^ 2 == 0.02 ", " (n == 861)"), parse = TRUE, col = "black", hjust = 0) +
  annotate(geom="text", x=2600, y=0.15, label="This study", color="black", hjust = 0) -> 
  factor_test1

factor_test1


allData2 %>% dplyr::filter(Rs_annual > 5000)
```


```{r}
allData2 %>% 
  mutate(Partition_group = case_when(
    Partition_method %in% c("Comparison", "Exclusion", "Girdling") ~ "Exclusion",
    Partition_method %in% c("Isotope") ~ "Isotope",
    Partition_method %in% c("Extraction") ~ "Physical seperation",
    Partition_method %in% c("Model", "TBCA", "Subtraction") ~ "Component",
    TRUE ~ "Other")) ->
  allData2

allData2 %>% 
  dplyr::select(RC_annual, Rs_annual, Partition_group) %>% 
  group_by(Partition_group) %>% 
  summarise(RC_mean = mean(RC_annual), Rs_mean = mean(Rs_annual),
            RC_sd = sd(RC_annual), Rs_sd = sd(Rs_annual), obs = n()) %>% 
  ggplot(aes(Rs_mean, RC_mean, col = Partition_group)) +
  geom_point(alpha = 1, aes(size = obs), show.legend = FALSE) +
  geom_errorbar(aes(ymin = RC_mean - RC_sd, ymax = RC_mean + RC_sd)) +
  # stat_function(fun = function(x) { 0.082 + 0.05*log(x) }, col = 'pink', size = 1.5) +
  # geom_smooth(method = "lm", col = "black", se = FALSE) +
  scale_color_discrete (name = "Partition") +
  labs(x = element_blank(),
       y = element_blank()) ->
  factor_test2
  # + geom_errorbarh(aes(xmin = Rs_mean - Rs_sd, xmax = Rs_mean + Rs_sd)) 

allData2 %>% 
  dplyr::select(RC_annual, Rs_annual, Partition_group) %>% 
  group_by(Partition_group) %>% 
  summarise(RC_mean = mean(RC_annual), Rs_mean = mean(Rs_annual),
            RC_sd = sd(RC_annual), Rs_sd = sd(Rs_annual), obs = n()) -> partition_test
lm(RC_mean ~ Rs_mean, data = partition_test) %>% summary()

# test climate
allData2 %>% 
  dplyr::select(RC_annual, Rs_annual, MiddleClimate) %>% 
  group_by(MiddleClimate) %>% 
  summarise(RC_mean = mean(RC_annual), Rs_mean = mean(Rs_annual),
            RC_sd = sd(RC_annual), Rs_sd = sd(Rs_annual), obs = n()) %>% 
  ggplot(aes(Rs_mean, RC_mean, col = MiddleClimate)) +
  geom_point(alpha = 1, aes(size = obs), show.legend = FALSE) +
  geom_errorbar(aes(ymin = RC_mean - RC_sd, ymax = RC_mean + RC_sd)) +
  # stat_function(fun = function(x) { 0.082 + 0.05*log(x) }, col = 'pink', size = 1.5) +
  # geom_smooth(method = "lm", col = "black", se = FALSE) +
  scale_color_discrete (name = "Climate") +
  labs(x = expression(Annual~R[S]~(g~C~m^{-2}~yr^{-1})),
       y = expression(R[root]~contribution~to~R[S]~(RC))) ->
  factor_test3

# test climate with tropical removed
allData2 %>% 
  dplyr::select(RC_annual, Rs_annual, MiddleClimate) %>% 
  filter(MiddleClimate %!in% c("A")) %>% 
  group_by(MiddleClimate) %>% 
  summarise(RC_mean = mean(RC_annual), Rs_mean = mean(Rs_annual),
            RC_sd = sd(RC_annual), Rs_sd = sd(Rs_annual), obs = n()) %>% 
  ggplot(aes(Rs_mean, RC_mean)) +
  geom_point(alpha = 1, aes(size = obs, col = MiddleClimate)) +
  geom_errorbar(aes(ymin = RC_mean - RC_sd, ymax = RC_mean + RC_sd)) +
  # stat_function(fun = function(x) { 0.082 + 0.05*log(x) }, col = 'pink', size = 1.5) +
  # geom_smooth(method = "lm", col = "black", se = FALSE) +
  scale_color_discrete (name = "Climate") +
  labs(x = expression(Annual~R[S]~(g~C~m^{-2}~yr^{-1})),
       y = expression(R[root]~contribution~to~R[S]~(RC)))

# test ecosystem
allData2$Ecosystem_type %>% unique()
globalData$Ecosystem2 %>% unique()
allData2$Ecosystem_type[allData2$Ecosystem_type == "Orchard"] = "Plantation"

Anova(lm(RC_annual ~ Ecosystem_type, data = allData2), type="III") 
allData2 %>% 
  dplyr::select(RC_annual, Rs_annual, Ecosystem_type) %>% 
  filter(Ecosystem_type %!in% c("Urban")) %>%
  group_by(Ecosystem_type) %>% 
  summarise(RC_mean = mean(RC_annual), Rs_mean = mean(Rs_annual),
            RC_sd = sd(RC_annual), Rs_sd = sd(Rs_annual), obs = n()) %>% 
  ggplot(aes(Rs_mean, RC_mean, col = Ecosystem_type)) +
  geom_point(alpha = 1, aes(size = obs), show.legend = FALSE) +
  geom_errorbar(aes(ymin = RC_mean - RC_sd, ymax = RC_mean + RC_sd)) +
  # stat_function(fun = function(x) { 0.082 + 0.05*log(x) }, col = 'pink', size = 1.5) +
  # geom_smooth(method = "lm", col = "black", se = FALSE) +
  scale_color_discrete (name = "Ecosystem") +
  labs(x = expression(Annual~R[S]~(g~C~m^{-2}~yr^{-1})),
       y = element_blank()) ->
  factor_test4
```

```{r,  fig.width=12, fig.height=6}
(factor_test1 | factor_test2) / (factor_test3 | factor_test4) +
  plot_annotation(tag_levels = "a")
```



## Test the relationship between RC and environmental factors
```{r function for factor testing}
# https://community.rstudio.com/t/passing-df-column-name-to-function/37293/3
lm_climate_factor <- function(col_name){
  subset(allData, MiddleClimate == var_eco[i]) %>% 
    dplyr::select(RC_annual, {{col_name}}, MiddleClimate) ->
    sub_data
  
  colnames(sub_data) <- c("RC_annual", "Factor", "Climate")
   
  sub_data %>% group_by(Climate, Factor) %>%
    summarise(RC_annual = mean(RC_annual), obs = n()) ->
    sub_data
  
  # return(sub_data)
  lm(RC_annual ~ Factor, data = sub_data) -> lm_model
  
  tibble(Factor = col_name,
         Climate = sub_data$Climate[1],
         obs = nrow(sub_data),
         R2 = summary(lm_model)$r.squared,
         Intercept = summary(lm_model)$coefficients[1,1] %>% round(6),
         Slope = summary(lm_model)$coefficients[2,1] %>% round(6),
         p_slope = summary(lm_model)$coefficients[2,4] %>% round(6)) ->
    model_outputs
  
  return(model_outputs)
}
```

### scatter plot
```{r test environmental factor, fig.width=5, fig.height=9}
# Figure S2
# test factors by climate type
var_eco <- sort(unique(allData$MiddleClimate))
### test bd
for(i in 1:8) {
  lm_climate_factor("bd_soilgrids") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, bd_soilgrids, MiddleClimate) %>% 
  group_by(MiddleClimate, bd_soilgrids) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_bd
sub_bd %>% 
  ggplot(aes(x = bd_soilgrids, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  geom_smooth(data = sub_bd %>% filter(MiddleClimate %in% c("A")), method = "lm", fill = "skyblue") +
  geom_smooth(data = sub_bd %>% filter(MiddleClimate %in% c("E")), method = "lm", fill = "pink", col = "red") +
  labs(x = expression(Bulk~density~(g~cm^{-3})), 
       y = expression(R[root]~contribution~to~R[S]~(RC))) +
  theme(legend.position = "none") +
  facet_grid(rows = vars(MiddleClimate)) ->
  bd_panel

sub_bd %>% 
  ggplot(aes(x = bd_soilgrids, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(data = sub_bd, method = "lm") +
  labs(x = expression(Bulk~density~(g~cm^{-3})), 
       y = expression(R[root]~contribution~to~R[S]~(RC))) +
  theme(legend.position = "none") ->
  bd_panel2

### EVI
for(i in 1:8) {
  lm_climate_factor("EVI_mean") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, EVI_mean, MiddleClimate) %>% 
  group_by(MiddleClimate, EVI_mean) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_evi
sub_evi %>% 
  ggplot(aes(x = EVI_mean, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(method = "lm") +
  labs(x = expression(EVI), 
       y = element_blank()) +
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_grid(rows = vars(MiddleClimate)) ->
  evi_panel

sub_evi %>% 
  ggplot(aes(x = EVI_mean, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(method = "lm") +
  labs(x = expression(EVI), 
       y = element_blank()) +
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) ->
  evi_panel2

### map
for(i in 1:8) {
  lm_climate_factor("MAP") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, MAP, MiddleClimate) %>% 
  group_by(MiddleClimate, MAP) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_map
sub_map %>% 
  ggplot(aes(x = MAP, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  geom_smooth(data = sub_map %>% filter(MiddleClimate %in% c("A")), method = "lm", fill = "skyblue") +
  labs(x = expression(MAP~(mm)), 
       y = element_blank()) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_grid(rows = vars(MiddleClimate)) ->
  map_panel

sub_map %>% 
  ggplot(aes(x = MAP, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(data = sub_map, method = "lm") +
  labs(x = expression(MAP~(mm)), 
       y = element_blank()) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) ->
  map_panel2

# Figure S3
### SOC
for(i in 1:8) {
  lm_climate_factor("soc_soilgrids") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, soc_soilgrids, MiddleClimate) %>% 
  group_by(MiddleClimate, soc_soilgrids) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_soc
sub_soc %>% 
  ggplot(aes(x = soc_soilgrids, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  geom_smooth(data = sub_soc %>% filter(MiddleClimate %in% c("B", "Df")), method = "lm", fill = "pink", col = "red") +
  geom_smooth(data = sub_soc %>% filter(MiddleClimate %in% c("E")), method = "lm", fill = "skyblue") +
  labs(x = expression(SOC~(Mg~C~ha^{-1})), 
       y = expression(R[root]~contribution~to~R[S]~(RC))) +
  theme(legend.position = "none") +
  facet_grid(rows = vars(MiddleClimate)) ->
  soc_panel

sub_soc %>% 
  ggplot(aes(x = soc_soilgrids, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(data = sub_soc, method = "lm") +
  labs(x = expression(SOC~(Mg~C~ha^{-1})), 
       y = expression(R[root]~contribution~to~R[S]~(RC))) +
  theme(legend.position = "none") ->
  soc_panel2

### mat
for(i in 1:8) {
  lm_climate_factor("MAT") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, MAT, MiddleClimate) %>% 
  group_by(MiddleClimate, MAT) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_mat
sub_mat %>% 
  ggplot(aes(x = MAT, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(method = "lm") +
  labs(x = expression(MAT~(degree~C)), 
       y = element_blank()) +
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_grid(rows = vars(MiddleClimate)) ->
  mat_panel

sub_mat %>% 
  ggplot(aes(x = MAT, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(method = "lm") +
  labs(x = expression(MAT~(degree~C)), 
       y = element_blank()) +
  theme(legend.position = "none",
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) ->
  mat_panel2

### BMb
for(i in 1:8) {
  lm_climate_factor("BM_belowground") %>% print()
}
allData %>% 
  dplyr::select(RC_annual, BM_belowground, MiddleClimate) %>% 
  group_by(MiddleClimate, BM_belowground) %>% 
  summarise(RC_annual = mean(RC_annual), obs = n()) ->
  sub_bmb
sub_bmb %>% 
  ggplot(aes(x = BM_belowground, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  geom_smooth(data = sub_bmb %>% filter(MiddleClimate %in% c("E")), method = "lm", fill = "skyblue", col = "blue") +
  labs(x = expression(Belowground~biomass~(Mg~C~ha^{-1})), 
       y = element_blank()) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  facet_grid(rows = vars(MiddleClimate)) ->
  bmb_panel

sub_bmb %>% 
  ggplot(aes(x = BM_belowground, y = RC_annual)) +
  geom_point(aes(size = obs), shape = 16, alpha = 0.5) +
  # geom_smooth(data = sub_bmb, method = "lm") +
  labs(x = expression(Belowground~biomass~(Mg~C~ha^{-1})), 
       y = element_blank()) +
  theme(axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) ->
  bmb_panel2
```

### output results
```{r, fig.width=9, fig.height=9}
bd_panel | evi_panel | map_panel
soc_panel | mat_panel | bmb_panel
```

```{r}
lm(RC_annual ~ bd_soilgrids, data = sub_bd) %>% summary()
lm(RC_annual ~ EVI_mean, data = sub_evi) %>% summary()
lm(RC_annual ~ MAP, data = sub_map) %>% summary()
lm(RC_annual ~ soc_soilgrids, data = sub_soc) %>% summary()
lm(RC_annual ~ MAT, data = sub_mat) %>% summary()
lm(RC_annual ~ BM_belowground, data = sub_bmb) %>% summary()
```

### output results - all data
```{r Factors testing, fig.width=8, fig.height=6}
(bd_panel2 | evi_panel2 | map_panel2) /
(soc_panel2 | mat_panel2 | bmb_panel2)
```


```{r}
allData %>% ggplot(aes(x = gpp_fluxcom, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

allData %>% ggplot(aes(x = clay_soilgrids, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

lm(RC_annual ~ EVI_mean * MiddleClimate, data = allData) %>% summary()
lm(RC_annual ~ bd_soilgrids * MiddleClimate, data = allData) %>% summary()
lm(RC_annual ~ MAP * MiddleClimate, data = allData) %>% summary()
lm(RC_annual ~ MAT * MiddleClimate, data = allData) %>% summary()
lm(RC_annual ~ soc_soilgrids * MiddleClimate, data = allData) %>% summary()
```

## partial dependent plot
```{r pdp}
pdp::partial(rf, pred.var = "MAT", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(MAT~(degree~C)),
       y = element_blank()) -> p_mat

pdp::partial(rf, pred.var = "MAP", plot = TRUE,
        plot.engine = "ggplot2") +
  xlab(expression(MAP~(mm))) +
  ylab(expression(Predicted~RC)) -> p_map

pdp::partial(rf, pred.var = "AM_percent", plot = TRUE,
        plot.engine = "ggplot2") +
  xlab(expression(MAP~(mm))) +
  ylab(expression(Predicted~RC)) -> p_am

pdp::partial(rf, pred.var = "EM_percent", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(EcM~("%")),
       y = element_blank()) -> p_ecm

pdp::partial(rf, pred.var = "ER_percent", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(ErM~("%")),
       y = element_blank()) -> p_erm

pdp::partial(rf, pred.var = "NM_percent", plot = TRUE,
        plot.engine = "ggplot2") +
  xlab(expression(MAP~(mm))) +
  ylab(expression(Predicted~RC)) -> p_nm

pdp::partial(rf, pred.var = "BM_aboveground", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(BMa~(Mg~C~ha^{-1})),
       y = element_blank()) -> p_bma

pdp::partial(rf, pred.var = "BM_belowground", plot = TRUE,
        plot.engine = "ggplot2") +
  xlab(expression(MAP~(mm))) +
  ylab(expression(Predicted~RC)) -> p_bmb

pdp::partial(rf, pred.var = "N_dep_half_deg", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(N~deposition~(g~N~m^{-2}~yr^{-1})),
      y = element_blank()) -> p_Ndep

pdp::partial(rf, pred.var = "bd_soilgrids", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(BD~(g~cm^{-3})),
       y = element_blank()) -> p_bd

pdp::partial(rf, pred.var = "clay_soilgrids", plot = TRUE,
        plot.engine = "ggplot2") +
  xlab(expression(MAP~(mm))) +
  ylab(expression(Predicted~RC)) -> p_clay

pdp::partial(rf, pred.var = "soc_soilgrids", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(SOC~(Mg~C~ha^{-1})),
       y = element_blank()) -> p_soc

pdp::partial(rf, pred.var = "EVI_mean", plot = TRUE,
        plot.engine = "ggplot2") +
  labs(x = expression(EVI),
       y = element_blank()) -> p_evi

```

```{r, fig.width=8, fig.height=8}
(p_map | p_bd | p_evi) / (p_bmb | p_soc | p_mat) / (p_clay | p_Ndep | p_bma) / (p_am | p_ecm) / (p_nm | p_erm)
  # plot_annotation(tag_levels = "a")
```

```{r}
drake::readd(BMa) %>% image()
drake::readd(BMa_agg) %>% image()

drake::readd(soilbd) %>% image()
drake::readd(soilbd_agg) %>% image()
```

