---
title: "SRPartitioning"
author: "Jinshi"
date: "10/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(results = 'hide', message = TRUE, include = TRUE, echo = FALSE,
                      fig.height = 5, fig.width = 7)

DATA_DIR <- 'Data'
OUT_DIR <- 'outputs'
```


```{r loading package}
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
theme_set(theme_bw())
library(here)
library(hexbin)
library(cowplot)
library(SpatialEpi)
library(patchwork)
library(ggridges)
# Note raster masks 'tidyr::extract' and 'dplyr::select'
library(raster)
library(randomForest)
library(caret)
library(e1071)
source("functions.R")
```


```{r prepare data}
srdb_all <- drake::readd(srdb_all) %>% 
  mutate(MAT = coalesce(MAT, MAT_WC),
         MAP = coalesce(MAP, MAP_WC),
         absLatitude = abs(Latitude),
         MiddleClimate = as.factor(MiddleClimate)) %>% 
  filter(RC_annual < 1 & RC_annual > 0)

# ecosystem2 NA to NotReported
srdb_all$Ecosystem2 <- as.character(srdb_all$Ecosystem2)
srdb_all[is.na(srdb_all$Ecosystem2),]$Ecosystem2 <- "Not reported"
# Make sure there is no missing data
allData = srdb_all %>% 
  dplyr::select(RC_annual, MAT, MAP, AM_percent, EM_percent, ER_percent, NM_percent,
                BM_aboveground, BM_belowground, N_dep_1993,
                # MiddleClimate,
                Ecosystem2, Latitude, Longitude,
                bd_soilgrids, clay_soilgrids, soc_soilgrids, EVI_mean) %>% na.omit()
  
# Extract global climate space data
map_mat_global <- drake::readd(map_mat_global)

# prepare data for global RC_annual prediction
globalData <- drake::readd(globalData) %>% 
  na.omit() %>% 
  mutate(absLatitude = abs(Latitude)) ->
  globalData

counties <- drake::readd(counties)
```


```{r check NULL values}
srdb_all %>% 
  filter(is.na(MAP))

allData %>% dplyr::select(Ecosystem2) %>% unique()
```


```{r srdb vasialization}
ggplot(srdb_all, aes(x = Leaf_habit, fill = Leaf_habit)) +
  geom_bar() +
  theme_minimal()

srdb_all %>% 
  ggplot(aes(Rs_annual, Leaf_habit, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(Annnual~R[S]~(g~C~m^{-2}~yr^{-1}))) +
  ylab(element_blank()) +
  scale_y_discrete(labels= c("Not available", "Deciduous", "Evergreen", "Mixed"))

srdb_all %>% 
  ggplot(aes(RC_annual, Leaf_habit, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(R[C]~Annual)) +
  ylab(element_blank()) +
  scale_y_discrete(labels= c("Not available", "Deciduous", "Evergreen", "Mixed"))

```


## Pulling the Climate Data and Comparing the Distribution to that in the SRDB
The first is a plot of the global distribution of mean annual temperature (MAT) and precipitation (MAP). 
The second also includes the values from all SRDB sites, plotted on top of the first graph.
```{r Site map, cache = TRUE, fig.height = 7, fig.width = 8}
# Plot panel a - global map with RC_annualsites
ggplot(data = counties) + 
  geom_polygon(aes(x = long, y = lat, group = group),
               color = "white", fill = 'gray', alpha = 0.5) + 
  guides(fill=FALSE) +
  geom_point(data = allData,
             aes(x=Longitude, y=Latitude, col = Ecosystem2),
             alpha = 0.75,
             stroke = 0.75 ) +
  scale_x_continuous(name="Longitude", breaks=seq(-180,180, 60),labels = seq(-180,180, 60))+
  scale_y_continuous(limits = c(-60, 80),name="Latitude", breaks=seq(-60,80,30),labels = seq(-60,80,30)) +
  scale_size_continuous(name = "RC_annual") +
  scale_colour_discrete(name = element_blank(),
                        breaks=c("Agriculture", "Grassland", "Not reported", "Savanna", "Shrubland", "T&B FOR", "Tropic FOR", "Urban", "Wetland"),
                        labels=c("Agriculture (n=261)", "Grassland (n=140)", "Not reported (n=2)", "Savanna (n=71)", "Shrubland (n=21)",
                                 "T&B FOR (n=256)", "Tropic FOR (n=70)", "Urban (n=2)", "Wetland (n=59)")) +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) ->
  gmap_a

# This chunk pulls climate data from WorldClim2 and makes some graphs from the mean annual temperature and precipitation data
# Download worldclim data for precip and tmean if necessary, into w10/ folder
# global distribution of MAT and MAP with points from the SRDB sites
map_mat_global %>% 
  ggplot(aes(x = mat/10, y = map)) +
  geom_hex(bins = 100, na.rm = TRUE) +
  scale_fill_viridis_c(name = "Grid cells", begin = 0.85, end = 0) +
  geom_point(data = srdb_all, aes(x = MAT_WC, y = MAP_WC),
             color = "black", shape = 1, size = 1.5, na.rm = TRUE) +
  # theme(legend.position = c(0.2, 0.65)) +
  labs(x = "MAT (°C)", y = "MAP (mm)") ->
  gmap_b

gmap_a / gmap_b +
  plot_annotation(tag_levels = 'a')

allData %>% 
  count(Ecosystem2)

srdb_all %>% 
  count(Ecosystem2)
```


```{r check WorldClim2 data}
# Comparing MAT
ggplot(srdb_all, aes(x = MAT_WC, y = MAT)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MATs to SRDB MATs", x = "WorldClim2 MAT (°C)", y = "Reported SRDB MAT (°C)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)

# Comparing MAP
ggplot(srdb_all, aes(x = MAP_WC, y = MAP)) + 
  geom_point(size = 1.5, color = "chartreuse3") +
  theme_minimal() +
  labs(title = "Comparing WorldClim2 MAPs to SRDB MAPs", x = "WorldClim2 MAP (mm)", y = "Reported SRDB MAP (mm)") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1)

## Makes some scatterplots with the SRDB data
ggplot(srdb_all, aes(x = MAT_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()

ggplot(srdb_all, aes(x = MAP_WC, y = Rs_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 3000) + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAP_WC, y = Rh_annual)) +
  geom_point(color = "chartreuse3") +
  ylim(0, 2000) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = MAT_WC, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = MAP_WC, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = MAP_WC, col = Leaf_habit)) +
  geom_point() +
  theme_minimal()
```

```{r Micro data graphs}

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = AM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = EM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = ER_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

ggplot(srdb_all, aes(x = Longitude, y = Latitude, col = NM_percent)) + 
  geom_point(size = 1.5) + 
  lims(x = c(-180, 180), y = c(-50, 80), color = c(0, 100)) + 
  theme_minimal()

# None of these seem to show correlations with mean annual precipitation, with the exception of ER mycorrhizae.
ggplot(srdb_all, aes(x = MAT_WC, y = AM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = EM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = MAT_WC, y = NM_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

# Microriza vs MAP
ggplot(srdb_all, aes(x = MAP_WC, y = ER_percent)) + 
  geom_point(size = 1.5, color = "chartreuse3") + 
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = AM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = EM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = ER_percent, fill = Leaf_habit)) +
  geom_boxplot() +
  theme_minimal()

ggplot(srdb_all, aes(x = Leaf_habit, y = NM_percent, fill = Leaf_habit)) +
  geom_violin(draw_quantiles = c(.25, .5, .75)) +
  theme_minimal()
```


```{r BD, SOC and clay vs RC_annual}
srdb_all %>% ggplot(aes(x = MAT, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = MAP, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = bd_soilgrids, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = clay_soilgrids, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = soc_soilgrids, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = soc_soilgrids, y = Rs_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = EVI_mean, y = Rs_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()

srdb_all %>% ggplot(aes(x = EVI_mean, y = RC_annual)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal()
```


## RF modelling
### Improving the model
Here is a graph of the error vs the number of trees in the model and a graph of how the error varies with different values of mtry.
mtry is a parameter for this random forest function that sets how many variables are tried when making each node of a regression tree.
The number of trees and this mtry value are both input parameters used to produce the random forest model.
```{r RF tune}
# Tune with different mtry values
tuneRF(subset(allData, select = -c(RC_annual, Site_ID, Longitude, Latitude, Manipulation)), 
       allData[, 6],
       stepFactor = 2,
       plot = TRUE,
       ntreetry = 200, # Based on the plot above, 200 trees seems reasonable
       trace = TRUE,
       improve = .01)
```

First, I will use the default settings in the randomForest package to make a random forest model to predict the root contribution to Rs (RC). 
```{r create RF model for prediction}
# set.seed(191982)
## Use random forest machine learning to create a model to predict Rs partitioning
# Set seed to a specific value, for reproducability
# Make a random forest model
model <- function() {
  randomForest(RC_annual ~ MAT + MAP +
                 AM_percent + EM_percent + ER_percent + NM_percent + BM_aboveground + BM_belowground +
                 N_dep_1993 + MiddleClimate + Ecosystem2 + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI_mean, 
               data = allData,
               # %>% sample_frac(frac_samp)
               importance = TRUE, 
               proximity = TRUE,
               ntree = 200,
               mtry = 4,
               na.action = na.exclude) }  

model()
plot(model())
importance(model())
varImpPlot(model())
```


```{r RF diagnose, fig.width=8, fig.height=6}
set.seed(20201007)
rf <- randomForest(RC_annual ~ MAT + MAP +
                     AM_percent + EM_percent + ER_percent + NM_percent + BM_aboveground + BM_belowground +
                     N_dep_1993 + MiddleClimate + Ecosystem2 + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI_mean, 
                   data = allData,
                   # %>% sample_frac(frac_samp)
                   importance = TRUE, 
                   proximity = TRUE,
                   ntree = 200,
                   mtry = 4,
                   na.action = na.exclude)

varImpPlot(rf)

# panel a
qplot(1:200, rf$mse ) + geom_line() + theme_bw() +
  xlab ("Number of trees (n)") + ylab("MSE") +
  theme(axis.title.y = element_text(angle = 90, hjust = 0.5) ) ->
  p_a


# panel b
tibble(x = c("ER%", "NM%", "EM%", "AM%", "Eco", "Climate", "BD", "SOC", "N dep", "BMa", "Clay%", "MAP", "MAT", "BMb", "EVI"),
       y = sort(importance(rf, type = 1)[1:15])) %>% 
  ggplot(aes(x = reorder(x, y), y)) +
  geom_bar(stat = "identity", width = 0.75) +
  coord_flip() +
  ylab ("Change of MSE (%)") +
  theme(axis.title.y = element_blank(),
        panel.grid.minor = element_blank()) ->
  rf_b

# panel c
tibble(x = c("ER%", "NM%", "EM%", "AM%", "Eco", "Climate", "N dep", "Clay%", "SOC", "BD", "BMb", "BMa", "MAT", "MAP", "EVI"),
       y = sort(importance(rf, type = 2)[1:15])) %>% 
  ggplot(aes(x = reorder(x, y), y)) +
  geom_bar(stat = "identity", width = 0.75) +
  coord_flip() +
  ylab ("Change of node purity") +
  theme(axis.title.y = element_blank(),
        panel.grid.minor = element_blank()) ->
  rf_c

# panel_b <- plot_grid(rf_b, rf_c, ncol = 2,
#                      # labels = c("b", "c"),
#                      # hjust = c(-9.5, -9.5),
#                      # vjust = 2.75,
#                      rel_widths = c(1, 1))
# 
# print(plot_grid(rf_a, panel_b, nrow = 2, labels = c("a", "b")
#                 , hjust = -1.5, vjust = 0.75) )

p_a / (rf_b | rf_c) + plot_annotation(tag_levels = 'a')
```


## SRDB RC_annual prediction
```{r SRDB RC_annual prediction check}
## Investigate how well the model works at predicting training data RC
# Use this model to predict RC values for the training data and put it all in one dataframe

# Plot training data RC values and the predicted values from the model
allData %>% 
  mutate(prediction = predict(model()),
         residuals = RC_annual - prediction) %>% 
  ggplot(aes(y = RC_annual, x = prediction)) +
  lims(x = c(0, 1)) +
  geom_point() +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1, color = "red")

allData %>% 
  mutate(prediction = predict(model()),
         residuals = RC_annual - prediction) %>%
  ggplot(aes(x = prediction, y = residuals)) +
  lims(x = c(0, 1)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed", size = 1, color = "red")
```


An mtry of 10 seem the best the most often in this test.
100 trees would probably be enough, but I will use 200 to be safe.
To make my predictions more consistent, I will make 20 separate models, using these values for the number of trees and mtry, and average the outputs.
```{r predict RC_annual of SRDB 20 times}
## Make 20 RF models with the better settings and average their predictions
tibble(pred_1 = predict(model(), allData),
       pred_2 = predict(model(), allData),
       pred_3 = predict(model(), allData),
       pred_4 = predict(model(), allData),
       pred_5 = predict(model(), allData),
       pred_6 = predict(model(), allData),
       pred_7 = predict(model(), allData),
       pred_8 = predict(model(), allData),
       pred_9 = predict(model(), allData),
       pred_10 = predict(model(), allData),
       pred_11 = predict(model(), allData),
       pred_12 = predict(model(), allData),
       pred_13 = predict(model(), allData),
       pred_14 = predict(model(), allData),
       pred_15 = predict(model(), allData),
       pred_16 = predict(model(), allData),
       pred_17 = predict(model(), allData),
       pred_18 = predict(model(), allData),
       pred_19 = predict(model(), allData),
       pred_20 = predict(model(), allData) ) ->
  RC_predict

RC_predict$RC_predict <- apply(RC_predict, 1, mean)  # applies function 'mean' to 1st dimension (rows)
```


## Global RC_annual prediction
```{r Predict global RC_annual}
# Predict global RC_annual
tibble(pred_1 = predict(model(), globalData),
       pred_2 = predict(model(), globalData),
       pred_3 = predict(model(), globalData),
       pred_4 = predict(model(), globalData),
       pred_5 = predict(model(), globalData),
       pred_6 = predict(model(), globalData),
       pred_7 = predict(model(), globalData),
       pred_8 = predict(model(), globalData),
       pred_9 = predict(model(), globalData),
       pred_10 = predict(model(), globalData),
       pred_11 = predict(model(), globalData),
       pred_12 = predict(model(), globalData),
       pred_13 = predict(model(), globalData),
       pred_14 = predict(model(), globalData),
       pred_15 = predict(model(), globalData),
       pred_16 = predict(model(), globalData),
       pred_17 = predict(model(), globalData),
       pred_18 = predict(model(), globalData),
       pred_19 = predict(model(), globalData),
       pred_20 = predict(model(), globalData) ) ->
  RC_predict_global

RC_predict_global$RC_predict <- apply(RC_predict_global, 1, mean)  # applies function 'mean' to 1st dimension (rows)


# Make the default theme have a centered title
theme_update(plot.title = element_text(hjust = 0.5))

# Plot predictions globally
cbind(globalData, RC_predict_global %>% dplyr::select(RC_predict)) %>% 
  ggplot(aes(Longitude, Latitude, color = RC_predict)) + 
  geom_point(size = 0.3) +
  lims(x = c(-180, 180), y = c(-60, 90)) +
  labs(color = "RC prediction") +
  scale_color_viridis_c()

# Look at the distribution of predictions
cbind(globalData, RC_predict_global %>% dplyr::select(RC_predict)) %>% 
  ggplot(aes(x = RC_predict)) +
  geom_histogram(color = "black", fill = "lightgrey") +
  labs(x = "RC Prediction")

# outputs RC_prediction
writ_file(cbind(globalData, RC_predict_global) %>%
            dplyr::select(Latitude, Longitude, RC_predict),
          "Global_RC_annual_prediction.csv")
```


```{r check N_dep}
ggplot(globalData, aes(Longitude, Latitude, color = EVI_mean)) + 
  geom_point(size = 0.3) +
  lims(x = c(-180, 180), y = c(-90, 90)) +
  labs(color = "RC prediction") +
  scale_color_viridis_c()

# N_dep_1993

# MAT + MAP + AM_percent + EM_percent + ER_percent + NM_percent + BM_aboveground + BM_belowground +
#   N_dep_1993 + MiddleClimate + Ecosystem2 + bd_soilgrids + clay_soilgrids + soc_soilgrids + EVI_mean
```


```{r predict srdb RC_annual}
allData = srdb_all %>% na.omit()
cbind(allData, RC_predict %>% dplyr::select(RC_predict)) %>% 
  mutate(reside = RC_annual - RC_predict) ->
  allData

allData %>% 
  ggplot(aes(y = RC_annual, x = RC_predict)) +
  lims(x = c(0, 1)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm") +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", size = 1, color = "red") +
  labs(y = expression(Measured~RC~annual),
       x = expression(Predicted~RC~annual)) +
  annotate("text", x = 0.1, y = 0.9, label = "italic(R) ^ 2 == 0.80", parse = TRUE) +
  scale_y_continuous(position = "left") ->
  srdb_pred_vs_measure

allData %>% 
  ggplot(aes(x = RC_predict, y = reside)) +
  lims(x = c(0, 1)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 0, linetype = "dashed", size = 1, color = "red")

lm(RC_annual ~ RC_predict, data = allData) %>% summary()
```


```{r RC_annual vs RC_prediction global, fig.width=8, fig.height=7}
globalData$RC_annual <- RC_predict_global$RC_predict

bind_rows(
  globalData %>% 
    mutate(Type = "I") %>% 
    dplyr::select(RC_annual, Type),
  
  allData %>% 
    mutate(Type = "III") %>% 
    dplyr::select(RC_annual, Type),
  
  allData %>% 
    mutate(Type = "II") %>% 
    dplyr::select(RC_predict, Type) %>% 
    dplyr::rename(RC_annual = RC_predict)
  ) %>% 
  ggplot(aes(RC_annual, Type, fill = 0.5 - abs(0.5 - stat(ecdf)))) +
  stat_density_ridges(geom = "density_ridges_gradient", calc_ecdf = TRUE) +
  scale_fill_viridis_c(name = "Probability", direction = -1) +
  xlab(expression(RC~Annual)) +
  ylab(element_blank()) +
  theme(axis.text.y = element_text(angle = 0)) +
  scale_y_discrete(labels= c("III", "II", "I"),
                   position = "left") ->
  overall_comparisom

globalData$RC_annual %>% min()
globalData$RC_annual %>% max()
globalData$RC_annual %>% mean()

allData$RC_annual %>% min()
allData$RC_annual %>% max()
allData$RC_annual %>% mean()

srdb_pred_vs_measure / overall_comparisom +  plot_annotation(tag_levels = 'a')
```

```{r}
# do local corelation and get coefficiency
corLocal(x, y, ngb=5, 
     method=c("pearson", "kendall", "spearman"), test=FALSE, filename='', ...)
```


